---
title: "Basic RevoScaleR Function"
author: "Redzuan"
date: "5/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

RevoScaleR is a R package developed by Microsoft that allow R data processing computation to be done in remote server without the need to pull the data into R environment in local machine. The remote computation context by RevoScaleR package can be utilized on SQL server (2016 onward) and distributed platforms such as Spark and Hadoop. Details description of this package can be found in Microsoft online document [here](https://docs.microsoft.com/en-us/machine-learning-server/r-reference/revoscaler/revoscaler).

In this post, basic use of RevoScaleR functions with data computation context in SQL Server will be demonstrated. They include:

+ Load CSV data to SQL Server database
+ Review data summary
+ Plot data
+ Import data from SQL to local R environment



## Setup and Prerequisite

This exercise was implemented using following environment and setup:

+ Microsoft SQL Server 2019 (RTM) - 15.0.2000.5 (X64) Developer Edition (64-bit) on Windows 10 Pro 10.0 with Machine Learning Services (R) installed
+ R version 3.5.2 (2018-12-20)
+ Run in local machine
+ Enable external script execution in SQL Server
+ Database permission to read, write and execute DDL permission is granted
+ sample_data Database was created in SQL server
+ Data input sample is from CSV format (Social_Network_Ads)


### Load CSV Data to SQL Server
```{r}

#Define SQL connection string 
sql_con_string <- "Driver=SQL Server;Server=DESKTOP-SKVB2GF;Database=sample_data;Trusted_Connection={Yes}"


#Generate data source object as S4 RxTextData class
social_netCSV <- RxTextData("../sample_data/Social_Network_Ads.csv")


#Define column type to get data type conversion to SQL format correctly

social_netCSV_colInfo <- list(`User ID` = list(type = "numeric"),
                              Gender = list(type = "factor", levels = c("Male", "Female")),
                              Age = list(type = "integer"),
                              EstimatedSalary = list(type = "numeric"),
                              Purchased = list(type = "factor", levels = c("0", "1")))


#Generate SQL server data source object to S4 class RxSqlServerData
social_netSQL <- RxSqlServerData(connectionString = sql_con_string,
                                 table = "social_network_ads",  #name of table in database
                                 colInfo = social_netCSV_colInfo)

#Load CSV data to database
rxDataStep(inData = social_netCSV,
           outFile = social_netSQL,
           overwrite = TRUE)


```



### Review Data Summary

RevoScaleR object can be summarized using ``rxSummary`` function. The formula describe in ``rxFormula`` can be passed to this syntax and it allow basic categorical variables manipulation. Refer ``?rxFormula`` for details


```{r}

#summary for all variables
rxSummary(~., data = social_netSQL) 

#Purchased count by Gender
rxSummary(~Purchased:Gender, data = social_netSQL) 

#Purchased summary by Age's statistic
rxSummary(~Purchased:Age, data = social_netSQL) 

#Purchased summary by group of Age
rxSummary(~Purchased:F(Age, low = 20, high = 25), data = social_netSQL) 

```


### Plot Data

```{r}
#Histogram
rxHistogram(~Age, data = social_netSQL)

rxHistogram(~Purchased | F(Age), data = social_netSQL)


#Line plot
rxLinePlot(Age ~ EstimatedSalary | Purchased, data = social_netSQL)


#Scatter plot
rxLinePlot(Age ~ EstimatedSalary | Purchased, data = social_netSQL, type = "p")


```


### Import Data from SQL to R Environment

```{r}
#Create SQL server data source object with SQL query

sql_QueryObject <- RxSqlServerData(sqlQuery = "SELECT Gender, Purchased, COUNT(*) AS Count
  FROM [sample_data].[dbo].[social_network_ads]
  GROUP BY Gender, Purchased",
  connectionString = sql_con_string)

#Import data

newDF <- rxImport(sql_QueryObject)



```

