---
title: "Date and POSIXct Class in R"
author: "Redzuan"
date: "4/4/2020"
output: html_document
---

Date and POSIXct, What are the main different.

 + Date and POSIXct are different class object. Date representing calendar date while POSIXct represent calendar dates and time
 + Date is stored as number of days from 1st Jan 1970
 + POSIXct is stored as number of seconds from 1st Jan 1970
 
 
The basis for any calculation and manipulation involving Date's and POSIXct object should considered above as main rules.

Let's see how both class behave in certain operation.


## Date class
```{r}
#Date class example produced from Sys.Date

date_class_ex <- Sys.Date()
date_class_ex

class(date_class_ex)


#Use unclass() to see what is a value that R store. This returns number of days since 1st Jan 1970

number_of_days_from_origin <- unclass(date_class_ex) 
number_of_days_from_origin


#Calculation for date class can be straighforward as simple numerical operation. The unit for number should be treat as day.

#Let's try with decrement date operation
date_of_origin <- date_class_ex - number_of_days_from_origin
date_of_origin





```


## POSIXct Class
```{r}
#POSIXct class object example

POSIXct_class_ex <- Sys.time()

POSIXct_class_ex

#class of POSIXct object
class(POSIXct_class_ex)


#Use unclass() to see what is a value that R store. This returns number of seconds since 1st Jan 1970

number_of_seconds_from_origin <- unclass(POSIXct_class_ex)
number_of_seconds_from_origin


#Calculation for POSIXct class can be straighforward as simple numerical operation. The unit for number should be treat as second.

#Let's try with decrement operation for POSIXct object

POSIXct_of_origin <- POSIXct_class_ex - number_of_seconds_from_origin
POSIXct_of_origin



```

 